Revisiting Multi-Domain Machine Translation

### 2020-04-15 2020/coling2020/multidomain.tex
full rewrite of sec 2, additional results and comments in section 4

\newcommand{\src}{\ensuremath{\mathbf{f}}} % source sentence
\newcommand{\trg}{\ensuremath{\mathbf{e}}} % target sentence
\newcommand{\domain}[1]{\texttt{\textsc{#1}}}
\newcommand{\system}[1]{\texttt{\textbf{#1}}}
\newcommand{\vlambda}{\ensuremath{\boldsymbol\lambda}\xspace} % parameters vector for a distribution

@@

Data-based Machine Translation (MT), whether statistical or neural, rests on well understood machine learning principles. Given a training sample of matched source-target sentence pairs $(\src,\trg)$\jcDone{shoulnt trg be e as below ?} drawn from an underlying distribution $\mathcal{D}$, a model parameterized by $\theta$ (here a translation function $f_{\theta}$) is trained by minimizing the empirical expectation of a loss function $\ell(f_\theta(\src), \trg)$. This approach ensures that the expectation of the translation loss will remain low when translating new sentences drawn from the same distribution.

Owing to the great variability of language data, this ideal situation is rarely met in practice, warranting the study of an alternative scenario, where the test distribution $\mathcal{D'}$ differs from $\mathcal{D}$. In this setting, \emph{domain adaptation} (DA) methods are required. DA has a long history in Machine Learning in general (eg.\ \cite{Shimodaira00improving,Ben-David09atheory})\fyTodo{cite non covariate shift} and in NLP in particular (eg.\ \cite{Daume06domain,Blitzer07domain,Jiang07instance}), and various techniques exist to handle both the situations where a (small) training sample drawn from $\mathcal{D'}$ is available in training, or where only samples of source-side (or target-side) sentences are available (see \cite{Foster07mixture,Bertoldi09domain,Axelrod11domain} for proposals from the statistical MT era, or \cite{Chu18asurvey} for a recent survey of DA in the context of Neural MT).
% In a nutshell, DA attempts to compensate for the distribution mismatch either by biasing the training sample to match ${\mathcal{D'}$ (eg.\ using resampling or instance-weighting strategies), or by adapting the

A seemingly related problem, that has important practical implications, is multi-domain (MD) machine translation \cite{Hassan17neural,Farajian17multidomain,Kobus17domaincontrol,Zeng18multidomain,Pham19generic} where one single system has to be trained and tested with data from multiple domains. As pointed out both by \newcite{Dredze08online} and \cite{Finkel09hierarchical}, this setting borrows much from DA, but also from multi-task learning \cite{Caruana97multitask,Yang15unified}; note that it also share commonalties with ensemble learning \cite{XXX}.\fyTodo{Missing ref} If the intuitions behind MDMT are simple, the exact specifications of MDMT systems are rarely spelled out: for instance, should they perform well when the test data is distributed like the train data, when it is equally distributed across training domains or when the test distribution is unknown? Should MDMT also be robust to new domains? How should they handle domain labeling errors? etc. % , which makes their evaluation fragile

@@

In this paper, we try to answer these questions and to spell out in a more precise fashion the requirements that a MDMT system should meet. Our main contribution is of methodological nature and consists of lists of expected properties of MDMT system and associated measurements to evaluate them. \fyTodo{New problems - continuous learning, automatic domains}Our second contribution is experimental, and we perform a thorough reanalysis of most recent multi-domain approaches from the literature, including a variant of a model initially introduced for DA. We show that\fyTodo{Spell out conclusions}.

@@

\subsection{Formalizing multi-domain learning \label{ssec:formalization}}

@@

\subsection{MDMT: a tentative definition \label{ssec:definition}}
Informally, a multi-domain system should be able to accurately translate heterogeneous sources of data. This loose definition correspond to several practical situations:
\begin{itemize}
\item the more common situation is when both train and test documents (or sentences) are domain-tagged, where each domain corresponds to a specific source of data, and differ from the other sources in terms of textual genre, thematic content \cite{Zhang16topicinformed}, register, style \cite{Niu18multitask}, etc. It is usually assumed that the set of test domains is the same as the set of training domains -- when this is not the case, we would have zero-shot domain adaptation, similar to what is studied in multilingual MT \cite{Firat16multiway,Ha16towards,Johnson17google,Platanios18contextual}. This is the more general setting, already proposed in \cite{Dredze09multidomain}.\fyTodo{do we need a name here: such as supervised MDMT ? and mostly supervised MDMT for the second case ?}

\item It can also be the case that domain labels are only known in training, while test inputs may or may not be associated with explicit domains. This is, for instance, the setting adopted \cite{Zeng18multidomain}, where automatic domain identification is performed for test sentences. Note that automatic domain identification can also be effectively performed in the first setting, especially in the presence of close or related domains in training.

\item A last setting, that has not received much attention for MT (see however \cite{Sennrich13multidomain}), is where both train and test domains are automatically infered from data, using clustering techniques. Even though domain labels are typically available in most training scenario, these approaches assumes that might be preferable, owing to the multiple sources of intra-corpus and inter-corpus variabilities, to resort to automatic domain assignments. The online approach of \newcite{Farajian17multidomain}, where distributional cues, rather that categorical tags, are used for fine-tuning, is a variant of this strategy.
\end{itemize}

In the rest of the paper, we mostly adhere to the first definition which correspond to a situation of practical interest for the industry, and assume that multi-domain MT is mostly about effectively using domain tags in training and testing. We also experiment in Section~\ref{sec:results} with the other two alternative scenarios, as they also define useful contexts to evaluate the actual robustness of MDMT systems.

\jcDone{...a very common scenario in MT industry where clients dispose of mulitiple data corresponding to different domains (differing in genre, thematic, register, style, etc.). }

\subsection{Reasons for using MDMT? \label{ssec:whymdmt}}
The motivations for considering MDMT are many. Looking at the literature, the main arguments in support of MDMT seem to be the following:
\begin{enumerate}
\item[[M1\hspace{-0.5em}]] computational / practical arguments. A first motivation for MDMT is practical  \cite{Sennrich13multidomain,Farajian17neural}: when faced with inputs that are potentially from multiple domains, it is easier and computationally cheaper to develop only one system instead of having to optimize (and maintain) multiple MT engines. Such arguments are all the more convincing that the number of domains considered is large, which is not always the case: for instance \cite{Britz17mixing,Zeng18multidomain,Jiang19multidomain} only consider 2-4 domains in their experiments.

\item[[M2\hspace{-0.5em}]] statistical arguments. The training data available for each domain is very unevenly distributed, and domain trained or adapted on small datasets will yield large estimator variance. Building mix-domain systems is likely to reduce this variance, at the expense of a larger statistical bias \cite{Clark12onesystem}. Under this view, MDMT would be especially useful for domains with little training data. This is what is observed for multilingual MT from English: an increase for under-resource languages due to positive transfer, at the cost of a decrease for well-resourced languages \cite{Arivazhagan19massively}. A slightly distinct line of reasoning is that mixing domains has a positive regularization effect for all domains. By introducing variability in training, it prevents domain adaptation from overfitting the available adaptation data and could help improve generalization even for well-resourced domains. A related case is made in \cite{Joshi12multidomain},
  where the benefits of MD are attributed to an ensembling effect, where several MT systems are simultaneously used in decoding; this effect may subsist even in the absence of clear domain separations.

\item[[M3\hspace{-0.5em}]] linguistic arguments. A third line of reasoning contends that domain specificities are mostly expressed lexically and will primarily affect content words or multi-word expressions; function words, on the other hand, are domain agnostic and remain semantically stable accross domain. A MDMT system should simultaneously learn lexical domain peculiarities, and leverage cross-domain similarities to improve the representation of generic contexts and words \cite{Zeng18multidomain,Pham19generic}. It is here expected that the MDMT scenario should be more profitable when the domain mix includes domains that are closely related and which can share more information. \fyTodo{polysemy}
\end{enumerate}

\fyTodo{recap}
These various motivations for adopting MDMT are not mutually exclusive. Each however yields specific expectations with respect to the performance of this approach: if the motivation is primarily computational, then a drop in MT quality with respect to fine-tuning each domain might be acceptable; if it is to improve statistical estimation, then the hope will be that MDMT will improve, at least for some domains, if not all, over individually trained systems.

@@

\subsection{Evaluation issues \label{ssec:evaluation}}

In theory, evaluation procedures should report realistic estimates of the test loss, computed on a sample deemed representative of the anticipated test conditions. Three scenarios need be distinguished:
\begin{itemize}
\item test data are mixed in the same proportion as training data; this warrants the design of an unbalanced mix-domain test sets in the same proportions as in training; alternatively, if per domain scores are ever reported, their average should take the mixing proportions into account;
\item all domains are equally likely and important in testing (the prior over test domains is uniform), which would warrant the computation of per domain scores and of unweighted averages, as reported in \cite{Farajian17multidomain,Su19exploring}. Note that in this situation, the optimal training regime would be to balance the training set in the same way -- which is hardly ever done in practice, as the baseline scores always consider mixing unevenly all available data. \cite{Wang20general} uses an intermediate mixing regime, which tends to reduce part of the unbalance, and improve the performance for underrepresented domains.
\item the test distribution is neither equal to the training set, nor uniform. Here again, if this distribution is known in advance, an optimal domain mix should be used in training to compute baseline scores, rather than using the empirical distribution of all the training data. A more challenging case is when the test set includes data from domains unseen in training, or when it is chosen adversarially \cite{XXX}.
% : this situation, where test instances have a null probability in training, goes agains the basic statistical assumptions of DA \cite{Shimodaira00improving}. It is thus important that the likeliness of such events should be clearly formulated when describing the evaluation protocol.
\end{itemize}

The papers in the literature are seldom explicit regarding their assumptions regarding the test distribution. In most cases, per-domain performances are reported, and even sometimes averaged, suggesting that the scenario~2 most faithfully reflects the expected testing conditions. Another important issue is the choice of the evaluation metrics: if BLEU is still the standard metric used in most studies, it is not necessary the best choice to measure subtle improvements of lexical choices \cite{Irvine13measuring} that are often associated with adapted systems.

A last issue regarding evaluation is related to the measurement of computational costs: when this is the main a motivation for using MDMT, we expect that computational costs would be reported and compared, which however is hardly ever done in practice. \fyTodo{Keep this ?}

\section{Challenging multi-domain translation systems}
In this section, we discuss seven\fyTodo{check number} operational requirements that can be expected from an effective multi-domain system, and discuss ways to evaluate whether these requirements are actually met. % in practice.

\subsection{Multi-domain systems should be effective \label{ssec:effective}}
% \subsubsection{Performance}
The first set of requirements for MDMT systems is that they should perform well in the face of mix-domain data. These yields the following requirements.

\paragraph{[P1-LAB]}\fyTodo{Decide naming scheme} A first expectation for MDMT is that they perform better that the baseline which disregards domain labels, or reassigns them in a random fashion \cite{Joshi12multidomain}.\fyTodo{Try random label assignments ?} % Testing that MDMT actually benefit from domain tags is simple, notwithstanding the previous remarks above regarding evaluation procedures.

\paragraph{[P2-TUN]} Additionally, following motivation~(M2) above, one would expect that MDMT improves over fine-tuning, at least in domains where data is scarce, or in situations where several domains are close. To evaluate this, we perform two measurements, using a real as well as an artificial scenario. In the real scenario, we simply compare the performance of MDMT and fine-tuning for domains of varying sizes, expecting a larger gain for smaller domains (requirement [P2.1-TUN]). In the artificial scenario, we split a single domain in two parts which are considered as distinct in training. The expectation here is that a MDMT should yield a clear gain for both pseudo sub-domains, which should benefit from the supplementary amount of relevant training (requirement [P2.2-TUN]). In this situation, MDMT should even outperform fine-tuning on either of the pseudo sub-domain.
\jcDone{I dont understand the prediction... why running MDMT on 2 artificial subdomains should show  gains over fine-tunning?}\fyTodo{small loss with respect to non split ?}

\subsection{Robustness with respect to domain definition \label{ssec:robusness}}
A second set of requirements is related to the definition of a domain. As repeatedly pointed out in the literature, training corpora for MT are often collected opportunistically and the view that each corpora constitutes a single domain is often a gross approximation.\footnote{Two of our own domains are actually composed of several subcorpora (IT and MED), see details in section~\ref{ssec:corpora}.} MDMT should aim to make the best of the available data and be robust to domain assignments. To challenge these requirements we propose the evaluate the following requirements.

\paragraph{[P3-HET]}
The notion of a domain is often a fragile one and an effective MDMT system should be able to discover not only when cross-domain sharing is useful (cf.\ property [P2.2-TUN])\jcDone{what is P2.2?}, but also when intra-domain heterogeneity is hurting. This requirement is tested by artificially conjoining two separate domains into one during training. The loss in performance with respect to the baseline (correct domain tags) should then be as small as possible.

@@

\paragraph{[P5-UNK]}
A related situation occurs when the domain of a test document is unknown. Several alternatives are in order: (a) automatically assigning one of the domain tag for decoding should not be much worse than using the true domain tag, assuming this domain is seen in training. For test documents from other domains, the same procedure can apply, and should outperform the use of a single mix-domain system.\fyTodo{Systems react to unknown domains}These requirements are respectively labelled [P5.1] and [P5.2]\fyTodo{Not so sure of these ones}.

\paragraph{[P6-DYN]}
Another requirement, more of an operational nature, is that a MDMT system should smoothly evolve to accomodate a larger number of domains, without having to retrain the full system each time new data is available. This is a requirement [P6] that we challenge by dynamically changing the number of training and test domains.

@@

\paragraph{[P7-NUM]} As mentioned above, the computational argument (M1) is all the more sensible as the number of domains increases, making the optimization of many individual systems both ineffective and undesirable. For lack of having access to corpora containing a large (eg.\ in the order of 100-1000 domains), we experiment with automatically learned domains.\fyTodo{considering a varying number of clusters.}

@@

We randomly select in each corpus a development and a test set of 1000 lines, and keep the rest for training. Validation sets are used to chose the best model according to the average BLEU score computed with\fyTodo{A word about meta-parameter settings}. Significance tests are performed using the bootstrap resampling method of \newcite{Koehn04statistical}. We report significant differences at the level of $p=$ \fyTodo{Fix correct p value}.

@@

We measure the distance between domains using the $\mathcal{H}$-Divergence introduced in \cite{Ben-David09atheory}, which relates domain similarity to the test error of a domain discriminator: the larger the error, the closer the domains. Formally, given test sets of size $m$ for domains $A$ and $B$, and $h(x)$ a trained domain predictor, $\mathcal{H}(A,B)$ is computed as:
$$
\mathcal{H}(A,B) = 2(1 - [\frac{1}{m} \sum_{x:h(x) = B} \mathbb{I}( x \in A) + \frac{1}{m} \sum_{x: h(x) = A} \mathbb{I}(x \in B)]),
$$
where $\mathbb{I}$ is the indicator function. Our domain discriminator is a multilayer perceptron independently trained for each pair of domains with sentence representations derived via mean pooling from the source side representation of our generic Transformer model. The hidden layer has dimension XXX, and the network is trained on balanced sets of YYY instances for ZZZ epochs.\fyTodo{Inform the classifier details}.

@@

\begin{itemize}
\item even when trained from entirely scratch, very few systems are able to take advantage of the additional data for the 6 initial domains - this is observed on the averaged scores in Table~\ref{tab:warmrestart} with hardly surpass those in Table~\ref{tab:performance}
\item restart is computationally cheaper (\fyTodo{how much cheaper here ?}) but underperforms cold start for all systems, sometimes by a wide margin.
\item the loss is largest for the small domains - how about distance to \domain{news} ?
\end{itemize}
\fyTodo{Very close to fine tuning towards News, no ?}:

>>>>

\newcommand{\src}{\ensuremath{\mathbf{f}}} % source sentence
\newcommand{\trg}{\ensuremath{\mathbf{e}}} % target sentence
\newcommand{\domain}[1]{\texttt{\textsc{#1}}}
\newcommand{\system}[1]{\texttt{\textbf{#1}}}
\newcommand{\vlambda}{\ensuremath{\boldsymbol\lambda}\xspace} % parameters vector for a distribution
\newcommand{\indic}[1]{\ensuremath{\mathbb{I}(#1)}}

@@

Data-based Machine Translation (MT), whether statistical or neural, rests on well understood machine learning principles. Given a training sample of matched source-target sentence pairs $(\src,\trg)$\jcDone{shoulnt trg be e as below ?} drawn from an underlying distribution $\mathcal{D}_s$, a model parameterized by $\theta$ (here, a translation function $h_{\theta}$) is trained by minimizing the empirical expectation of a loss function $\ell(h_\theta(\src), \trg)$. This approach ensures that the translation loss remains low when translating more sentences drawn from the same distribution.

Owing to the great variability of language data, this ideal situation is rarely met in practice, warranting the study of an alternative scenario, where the test distribution $\mathcal{D}_t$ differs from $\mathcal{D}_s$. In this setting, \emph{domain adaptation} (DA) methods are in order. DA has a long history in Machine Learning in general (eg.\ \cite{Shimodaira00improving,Ben-David09atheory,Pan10asurvey})\fyDone{cite non covariate shift} and in NLP in particular (eg.\ \cite{Daume06domain,Blitzer07domain,Jiang07instance}) and various techniques exist to handle both the situations where a (small) training sample drawn from $\mathcal{D}_t$ is available in training, or where only samples of source-side (or target-side) sentences are available (see \cite{Foster07mixture,Bertoldi09domain,Axelrod11domain} for proposals from the statistical MT era, or \cite{Chu18asurvey} for a recent survey of DA for Neural MT).
% In a nutshell, DA attempts to compensate for the distribution mismatch either by biasing the training sample to match ${\mathcal{D'}$ (eg.\ using resampling or instance-weighting strategies), or by adapting the

A seemingly related problem is multi-domain (MD) machine translation \cite{Hassan17neural,Farajian17multidomain,Kobus17domaincontrol,Zeng18multidomain,Pham19generic} where one single system has to be trained as well as tested with data from multiple domains. MD machine translation corresponds to a very common situation, where all available data, no matter its origin, is used to train a system that will be robust to domain variability in testing.  
% As pointed out both by \newcite{Dredze08online} and \cite{Finkel09hierarchical}, this setting borrows much from DA, but also from multi-task learning \cite{Caruana97multitask,Yang15unified}.
If the intuitions behind MDMT are quite simple, the exact specifications of MDMT systems are rarely spelled out: for instance, should MDMT perform well when the test data is distributed like the training data, when it is equally distributed across domains or when the test distribution is unknown? Should MDMT also be robust to new domains? How should it handle domain labeling errors? % , which makes their evaluation fragile

@@

In this paper, we try to study these questions and to spell out in a more precise fashion the requirements that a MDMT system should meet. Our first contribution is of methodological nature and consists of lists of expected properties of MDMT system and associated measurements to evaluate them. In doing so, we also shed lights on new problems that arise in this context, regarding for instance the accomodation of new domains in the course of training, or the computation of automating domain tags.\fyDone{New problems - continuous learning, automatic domains} Our second main contribution is experimental, and consists in a thorough reanalysis of eight recent multi-domain approaches from the literature, including a variant of a model initially introduced for DA. We show that\fyTodo{Spell out conclusions}.

@@

\subsection{Complements - to be removed \label{ssec:oldformalization}}

@@

% ----
\subsection{Formalizing multi-domain translation \label{ssec:formalization}}

We conventionally define a domain $d$ as a distribution $\mathcal{D}_d(x)$ over some feature space $\mathcal{X}$ that is shared across domains \cite{Pan10asurvey}: in machine translation, $\mathcal{X}$ is the representation space for source sentences; each domain corresponds to a specific source of data, and differs from the other data sources in terms of textual genre, thematic content \cite{Zhang16topicinformed}, register, style \cite{Niu18multitask}, etc. Translation in domain $d$ is formalized by a translation function $h_d(y|x)$ pairing sentences in a source language with sentences in a target language $y \in \mathcal{Y}$. $h_d$ is usually assumed to be deterministic (hence $y = h_d(x)$, but can differ from one domain to the other.

A typical learning scenario in MT is to have access to a samples from $n_d$ domains, which means that the training distribution $\mathcal{D}^s$ is a mixture $\mathcal{D}^s(x) = \sum_d \lambda^{s}_{d} \mathcal{D}_d(x)$. Multi-domain learning, as defined in \cite{Dredze08online} further assumes that domain tags are also available in testing; the implication being that the test distribution is also as a mixture $\mathcal{D}^t(x) = \sum_d \lambda^{t}_{d} \mathcal{D}_d(x)$ of several domains, making the problem distinct from mere domain adaption.  A multi-domain learner is then expected to use these tags effectively \cite{Joshi12multidomain} when computing the combined translation function $h(x,d)$, and to perform well in all domains \cite{Finkel09hierarchical}. Note that this setting is closely related to the multi-source adaptation problem studied in \cite{Mansour09domainadaptation,Mansour09multiple}.

This definition seems to be the most accepted view of a multi-domain MT\footnote{One important exception is in \cite{Zeng18multidomain}, where test translations rely on a similarity score between sentences, rather than domain labels.}, and one that we also adopt here. Note that in the absence of further specification, the naive answer to the MD setting should be to train one distinct translation function $\hat{h}_d(x)$ for each domain, then translate using $\hat{h}(x,d) = \sum_{d'} h_{d'}(x) \indic{d' = d}$. We now discuss the arguments that are usually put forward to proceed differently.
% --

% This is also the view that we adopt here, where we assume training inputs in $(\mathcal{X}\times{}\{1\dots{}D\})$, sampled from a mixture $\sum \lambda_{s,d}\mathcal{D}_{d} \times \{d\}$, and also consider various test scenarios, depending whether $\mathcal{D}_t$ is a fixed mixture of known domains, include new domain(s), in the supervised and unsupervised, and zero-shot settings.\fyTodo{explain zero shot}.

\subsection{Reasons for building MDMT systems \label{ssec:whymdmt}}

A first motivation for moving away from the one-domain / one-system solution are practical  \cite{Sennrich13multidomain,Farajian17neural}: when faced with inputs that are potentially from multiple domains, it is easier and computationally cheaper to develop one single system instead of having to optimize and maintain multiple engines. The underlying assumption here is that the number of domains of interests can be large, a limiting scenario being fully personalized machine translation \cite{Michel2018extreme}.
% which is not always the case: for instance \cite{Britz17mixing,Zeng18multidomain,Jiang19multidomain} only consider 2-4 domains in their experiments.\fyTodo{Keep this last part ?}

A second line of reasoning rests on linguistic properties of the translation function and contends that domain specificities are mostly expressed lexically and will primarily affect content words or multi-word expressions; function words, on the other hand, are domain agnostic and tend remain semantically stable accross domains, motivating some cross-domain parameter sharing. A MDMT system should simultaneously learn lexical domain peculiarities, and leverage cross-domain similarities to improve the translation of generic contexts and words \cite{Zeng18multidomain,Pham19generic}. It is here expected that the MDMT scenario should be more profitable when the domain mix includes domains that are closely related and which can share more information.

A third series of motivations are of statistical nature. The training data available for each domain is usually unevenly distributed, and domain-specific systems trained or adapted on small datasets are likely to have a high variance and perform poorly. Training mix-domain systems is likely to reduce this variance, at the expense of a larger statistical bias \cite{Clark12onesystem}. Under this view, MDMT would be especially beneficial for domains with little training data. This is observed for multilingual MT from English: an increase for under-resource languages due to positive transfer, at the cost of a decrease in performance for well-resourced languages \cite{Arivazhagan19massively}.

Combining multiple domain-specific MTs can also been justified in the sake of distributional robustness \cite{Mansour09domainadaptation,Mansour09multiple}, for instance when the test mixture differs from the train mixture, or when it includes new domains unseen in training. Such scenario is already well documented for zero-shot multi-lingual MT \cite{Firat16multiway,Ha16towards,Johnson17google,Platanios18contextual}, where mixing languages has more than demonstrated its usefulness.  An even more challenging case is when the MT would need to perform well for any test distribution (as considered eg.\ in \cite{Hoffman18algorithms,Oren19distributionally}  for other tasks). In all these cases, mixing domains in training and/or testing is likely to improve robustness against unexpected inputs.

A distinct line of reasoning is that mixing domains can a positive regularization effect for all domains. By introducing variability in training, it prevents domain adaptation from overfitting the available adaptation data and could help improve generalization even for well-resourced domains. A related case is made in \cite{Joshi12multidomain}, where some of the benefits of MD training are attributed to an ensembling effect, where several systems from multiple domains are simultaneously used in the prediction phase\cite{Saunders19ucam}; this effect may subsist even in the absence of clear domain separations.

\fyDone{recap}
To recap, there are multiple arguments for adopting MDMT, some already used in DA settings, and some are new. These arguments are not mutually exclusive; however each yields specific expectations with respect to the performance of this approach, and should also yield appropriate evaluation procedure. If the motivation is primarily computational, then a drop in MT quality with respect to multiple individual domain might be acceptable if compensated by the computational savings. If it is to improve statistical estimation, then the hope will be that MDMT will improve, at least for some under-resourced domains, over individually trained systems. If finally, it is to make the system more robust to unexpected or adversarial test distributions, then this is the setting that should be used to evaluate MDMT. The next section discusses ways in which these various requireents of MDMT systems can ben challenged.

@@

% \subsection{MDMT: a tentative definition \label{ssec:definition}}

% Informally, a multi-domain system should be able to accurately translate heterogeneous sources of data. This loose definition correspond to several practical situations:
% \begin{itemize}
% \item the more common situation is when both train and test documents (or sentences) are domain-tagged, where each domain corresponds to a specific source of data, and differ from the other sources in terms of textual genre, thematic content \cite{Zhang16topicinformed}, register, style \cite{Niu18multitask}, etc. It is usually assumed that the set of test domains is the same as the set of training domains -- when this is not the case, we would have zero-shot domain adaptation, similar to what is studied in multilingual MT. This is the more general setting, already proposed in \cite{Dredze09multidomain}.\fyTodo{do we need a name here: such as supervised MDMT ? and mostly supervised MDMT for the second case ?}

% \item It can also be the case that domain labels are only known in training, while test inputs may or may not be associated with explicit domains. This is, for instance, the setting adopted \cite{Zeng18multidomain}, where automatic domain identification is performed for test sentences. Note that automatic domain identification can also be effectively performed in the first setting, especially in the presence of close or related domains in training.

% \item A last setting, that has not received much attention for MT (see however \cite{Sennrich13multidomain}), is where both train and test domains are automatically infered from data, using clustering techniques. Even though domain labels are typically available in most training scenario, these approaches assumes that might be preferable, owing to the multiple sources of intra-corpus and inter-corpus variabilities, to resort to automatic domain assignments. The online approach of \newcite{Farajian17multidomain}, where distributional cues, rather that categorical tags, are used for fine-tuning, is a variant of this strategy.
% \end{itemize}

% In the rest of the paper, we mostly adhere to the first definition which correspond to a situation of practical interest for the industry, and assume that multi-domain MT is mostly about effectively using domain tags in training and testing. We also experiment in Section~\ref{sec:results} with the other two alternative scenarios, as they also define useful contexts to evaluate the actual robustness of MDMT systems.

% \jcDone{...a very common scenario in MT industry where clients dispose of mulitiple data corresponding to different domains (differing in genre, thematic, register, style, etc.). }

% \subsection{Evaluation issues \label{ssec:evaluation}}

% In theory, evaluation procedures should report realistic estimates of the test loss, computed on a sample deemed representative of the anticipated test conditions. Three scenarios need be distinguished:
% \begin{itemize}
% \item test data are mixed in the same proportion as training data; this warrants the design of an unbalanced mix-domain test sets in the same proportions as in training; alternatively, if per domain scores are ever reported, their average should take the mixing proportions into account;
% \item all domains are equally likely and important in testing (the prior over test domains is uniform), which would warrant the computation of per domain scores and of unweighted averages, as reported in \cite{Farajian17multidomain,Su19exploring,Wang20general}. Note that in this situation, the optimal training regime would be to balance the training set in the same way -- which is hardly ever done in practice, as the baseline scores always consider mixing unevenly all available data. \cite{Wang20general} uses an intermediate mixing regime, which tends to reduce part of the unbalance, and improve the performance for underrepresented domains.
% \item the test distribution is neither equal to the training set, nor uniform. Here again, if this distribution is known in advance, an optimal domain mix should be used in training to compute baseline scores, rather than using the empirical distribution of all the training data. A more challenging case is when the test set includes data from domains unseen in training, or when it is chosen adversarially \cite{XXX}.
% % : this situation, where test instances have a null probability in training, goes agains the basic statistical assumptions of DA \cite{Shimodaira00improving}. It is thus important that the likeliness of such events should be clearly formulated when describing the evaluation protocol.
% \end{itemize}

% The papers in the literature are seldom explicit regarding their assumptions regarding the test distribution. In most cases, per-domain performances are reported, and even sometimes averaged, suggesting that the scenario~2 most faithfully reflects the expected testing conditions. Another important issue is the choice of the evaluation metrics: if BLEU is still the standard metric used in most studies, it is not necessary the best choice to measure subtle improvements of lexical choices \cite{Irvine13measuring} that are often associated with adapted systems.

% A last issue regarding evaluation is related to the measurement of computational costs: when this is the main a motivation for using MDMT, we expect that computational costs would be reported and compared, which however is hardly ever done in practice. \fyTodo{Keep this ?}

\section{Challenging multi-domain translation systems \label{sec:challenging}}
In this section, we discuss seven\fyTodo{check number} operational requirements that can be expected from an effective multi-domain system, and discuss ways to evaluate whether these requirements are actually met.\fyTodo{Insert discussion about scores}

\subsection{Multi-domain systems should be effective \label{ssec:effective}}
% \subsubsection{Performance}
A first expectation is that MDMT systems should perform well in the face of mixed-domain test data. This yields the following three requirements.

\paragraph{[P1-LAB]}\fyTodo{Decide naming scheme} A MDMT should perform better than the baseline which disregards domain labels, or reassigns them in a random fashion \cite{Joshi12multidomain}. Evaluating this requirement is a matter of a mere comparison, assuming the test distribution of domains is known: if all domains are equally important, performance averages can be reported; if they are not, weighted averages should be used instead.

\fyTodo{Try random label assignments ?} % Testing that MDMT actually benefit from domain tags is simple, notwithstanding the previous remarks above regarding evaluation procedures.

\paragraph{[P2-TUN]} Additionally, one can expect that MDMT will improve over fine-tuning \cite{Luong15stanford,Freitag16fast}, at least in domains where data is scarce, or in situations where several domains are close. To evaluate this, we perform two measurements, using a real as well as an artificial scenario. In the real scenario, we simply compare the performance of MDMT and fine-tuning for domains of varying sizes, expecting a larger gain for smaller domains. In the artificial scenario, we split a single domain in two parts which are considered as distinct in training. The expectation here is that a MDMT should yield a clear gain for both pseudo sub-domains, which should benefit from the supplementary amount of relevant training. In this situation, MDMT should even outperform fine-tuning on either of the pseudo sub-domain.
\jcDone{I dont understand the prediction... why running MDMT on 2 artificial subdomains should show  gains over fine-tunning?}\fyTodo{small loss with respect to non split ?}

\subsection{Robustness with respect to domain definition \label{ssec:robusness}}
A second set of requirements is related to the definition of a domain. As repeatedly pointed out in the literature, parallel corpora used MT are often collected opportunistically and the view that each corpus constitutes a single domain is often a gross approximation.\footnote{Two of our own ``domains'' actually comprise several subcorpora (IT and MED), see details in section~\ref{ssec:corpora}.} MDMT should aim to make the best of the available data and be robust to domain assignments. To challenge these requirements we propose the evaluate the following requirements.

\paragraph{[P3-HET]}
The notion of a domain being a fragile one, an effective MDMT system should be able to discover not only when cross-domain sharing is useful (cf.\ requirement [P2-TUN])\jcDone{what is P2.2?}, but also when intra-domain heterogeneity is hurting. This requirement is tested by artificially conjoining two separate domains into one during training. The loss in performance with respect to the baseline (using correct domain tags) should then be as small as possible.

@@

\paragraph{[P5-UNK]}
A related situation occurs when the domain of a test document is unknown. Several situations need be considered: when the domain is seen in training, then using automatically predicted domain labels should not be much worse than using the correct one. For test documents from other domains, the same procedure can be used, and should outperform the use of a single mixed-domain system.\fyTodo{Systems react to unknown domains}
% These requirements are respectively labelled [P5.1] and [P5.2]\fyTodo{Not so sure of these ones}.

\paragraph{[P6-DYN]}
Another requirement, more of an operational nature, is that a MDMT system should smoothly evolve to accomodate a growing number of domains, without having to retrain the full system each time new data is available. This is a requirement [P6-DYN] that we challenge by dynamically changing the number of training and test domains.

@@

\paragraph{[P7-NUM]} As mentioned above, MDMT systems have often been motivated by computational arguments. This argument is all the more sensible as the number of domains increases, making the optimization of many individual systems both ineffective and undesirable. For lack of having access to corpora containing very large sets (eg.\ in the order of 100-1000) domains, we experiment with automatically learned domains.\fyTodo{considering a varying number of clusters.}

@@

We randomly select in each corpus a development and a test set of 1,000 lines and keep the rest for training. Validation sets are used to chose the best model according to the average BLEU score \cite{Papineni02bleu}.\fyDone{A word about meta-parameter settings}, computed with\fyTodo{}. Significance tests are performed using the bootstrap resampling method of \newcite{Koehn04statistical} and implementation of compare-mt\footnote{\url{https://github.com/neulab/compare-mt}} \cite{Neubig19compare-mt}. We report significant differences at the level of $p=0.05$\fyDone{Fix correct p value}.

@@

We measure the distance between domains using the $\mathcal{H}$-Divergence introduced in \cite{Ben-David09atheory}, which relates domain similarity to the test error of a domain discriminator: the larger the error, the closer the domains. Formally, given test sets of size $m$ for domains $A$ and $B$, and $h(x)$ a trained domain predictor, $\mathcal{H}(A,B)$ is computed as:
$$
\mathcal{H}(A,B) = 2(1 - [\frac{1}{m} \sum_{x:h(x) = B} \mathbb{I}( x \in A) + \frac{1}{m} \sum_{x: h(x) = A} \mathbb{I}(x \in B)]),
$$
where $\mathbb{I}$ is the indicator function. Our domain discriminator is SVM independently trained for each pair of domains with sentence representations derived via mean pooling from the source side representation of our generic Transformer model. We used the sci-kit learn implementation with default values.\fyDone{Inform the classifier details}\fyDone{Insert tableau} Results in Table~\ref{tab:domaindist} show that all domains are well defined and separated from the all others, with \domain{rel} being the furthest apart, while \domain{talk} is slightly more central.

\begin{table}\centering
  \begin{tabular}{|l*{5}{|r}|} \hline
    & law & bank & talk & IT & rel \\ \hline
    med &1.93 &1.97 &1.9 &1.93 &1.97 \\
    law   && 1.94 & 1.97 &1.93 & 1.99 \\
    bank &&&1.98 &1.94 &1.99 \\
    talk   &&&&1.92 &1.93 \\
     IT     &&&&& 1.99 \\ \hline
  \end{tabular}
  \caption{The $\mathcal{H}$-divergence between domains}
  \label{tab:domaindist}
\end{table}

@@

\begin{itemize}
\item even when trained from entirely scratch, very few systems are able to take advantage of the additional data for the 6 initial domains - this is observed on the averaged scores in Table~\ref{tab:warmrestart} with hardly surpass those in Table~\ref{tab:performance}
\item restart is computationally cheaper (\fyTodo{how much cheaper here ?}) but underperforms cold start for all systems, sometimes by a wide margin.
\item the loss is largest for the small domains - how about distance to \domain{news} ?
\end{itemize}
\fyDone{Very close to fine tuning towards News, no ?}
